### 2016-12-01 dissertation/intro.tex
dissertation v2_chap 3 4_update1
dissertation v2: part intro update 1

\chapter{Introduction} \label{ch:intro}
There is an ever-increasing need for secure communication within information 
technology. Security of sensitive information relies more and more heavily 
on encryption methodologies implemented in hardware by cryptographic 
circuits. One of the most prominent of these methodologies is Elliptical 
Curve Cryptography (\emph{ECC}), which provides more strength per encryption bit 
than other encryption methodologies.
The main building blocks of ECC hardware implementations 
are fast, \emph{custom-built Galois field arithmetic circuits}.
These circuits are notoriously hard to verify, yet their correctness is 
vitally important in critical applications. In \cite{crypto:bug_attacks}, 
for example, it is shown that a bug in the hardware could lead to the full 
leakage of the secret cryptographic key, which could compromise the entire
system. Thus, formal verification is imperative in Electronic Design 
Automation (\emph{EDA}) when dealing with cryptographic circuits.

To facilitate this verification, it is highly desirable to obtain a {\it word-level 
representation of the datapath of the ECC arithmetic block} from its 
bit-level implementation. Ideally, this abstraction should be {\it canonical},
as this allows the it to be directly applicable to equivalence checking. 
Such a canonical, word-level abstraction of the 
Galois field arithmetic block would not only make it easier to verify and 
reason 
about the cryptographic system as a whole, but also enable the use of 
higher level abstraction and synthesis tools. 
As arithmetic circuits are custom-designed, often modularly, using Galois field arithmetic blocks, 
the abstraction should also exploit the hierarchical nature of the circuitry.
Due to the modular circuit structure, abstraction of each arithmetic block
becomes the key in verification of the full circuit. 
Practical 
applications of ECC dictate a datapath of a minimum of $163$-bits, up to 
$571$-bits, as 
designated by the National Institute for Standards and Technology (NIST).
However abstraction of 
Galois field arithmetic circuits has been infeasible for data-paths beyond 
$16$ bits.

This dissertation proposes an algebraic geometry based approach to abstract canonical, 
word-level representations of bit-level Galois field arithmetic circuits.
The approach is able to abstract representations
for circuits up to 571 bits in size, which is the largest NIST
standard for datapath size in ECC. 
Verification of circuits for which this abstraction has been computed is shown to 
be trivial; thus, the focus is on deriving the abstraction quickly and efficiently.
%Furthermore, we show how we can use these abstract representations to 
%perform improved formal verification of these circuits.

\section{Hardware Design and Verification Overview}
The typical design flow of a hardware system, 
as shown in Fig. \ref{fig:cadflow}, starts with a hardware system 
specification, which describes the necessary functions and parameters that 
the system must perform and adhere to. The specification is typically
modelled using a transaction-level model (\emph{TLM}), which describes 
communication details between large circuit modules. 
The TLM is then translated 
into a register-transfer-level (\emph{RTL}) description, which is composed of abstracted, 
interconnected circuit blocks that compose the entire system. RTL is 
typically implemented 
in hardware description languages (\emph{HDL}) such as Verilog and VHDL, 
which are the most popular choices in the industry.
Next, the RTL is optimized and converted into a \emph{netlist}, i.e. a large 
collection of small physical blocks (MOSFET, Boolean logic gates, etc.) and 
the inter-connections (wires) between them. 
Lastly, the netlist is further optimized and then mapped onto a physical space 
on a chip, which is then sent off for fabrication. This entire design flow
is automated by Computer-Aided Design (\emph{CAD}) tools.

@@

\section{Dissertation Objective, Motivation, and Contributions}
This dissertation focuses on abstracting a canonical, 
word-level representation of hardware (bit-level) implementations of 
combinational circuits. The proposed technique is a full abstraction solution which can be 
applied to any arbitrary acyclic combinational circuit. 
It is particularly efficient when applied to Galois field arithmetic circuits.
Using this technique, if the abstraction of the circuit's implementation and its
specification are found, they can be easily compared to determine equivalence.
Implementation of a custom software tool, developed to compute the abstractions, is
also described.

\subsection{Motivating Application}
The motivation for this work comes from applications of Galois field 
arithmetic circuits in elliptical curve cryptography ({\it ECC}) hardware systems.
The main operations of encryption, decryption, and 
authentication in ECC rely on operations performed on elliptic curves, which 
are implemented in hardware as polynomial functions over Galois fields. 
To be applicable in real-world situations, ECC data-paths
should be a minimum of $163$-bits wide, which is the minimum NIST standard, 
up to a recommended size of $571$-bit operand widths. Many non-ECC cryptosystems
have datapaths on the order of $1000$-bits.

A Galois field arithmetic circuit with a datapath size of $k$ 
is built as a Boolean function: $\mathbb{B}^k \rightarrow \mathbb{B}^k$. 
This function is mapped to an operation 
$f:\mathbb{F}_{2^k} \rightarrow \mathbb{F}_{2^k}$ 
over the Galois field $\mathbb{F}_{2^k}$. 
These circuits are custom-built, modular
systems which cannot be synthesized due to their complex nature. Thus, 
formal verification is needed to ensure they operate correctly.

Recent computer-algebra based formal verification techniques have been
able to perform verification of Galois field arithmetic circuits with
a datapath size up to $163$-bits \cite{lv:phd}. Word-level abstractions 
of Galois
field arithmetic circuits could be used to further improve these formal 
verification techniques to allow for verification of larger circuits, as
well as provide the other benefits of word-level abstraction.
However, there is currently no technique for computing word-level 
abstractions of Galois field circuits of any practical size.

{
\begin{figure}[h]
\centerline{
\includegraphics[width=0.5\textwidth]{./figures/interpolate}
}
\caption{Circuit with $k$-bit input $A$ and $k$-bit output $Z$. 
Abstraction to be derived as $Z=\Func(A)$.}
\label{fig:abstractA_Z}
\end{figure}
}

While the motivation comes from the need to verify Galois field arithmetic
circuits, the presented approach can be generalized to be applicable to any
combinational acyclic circuit.
Any such circuit with a $k$-bit input $A$ and a $k$-bit output $Z$, such
as the one shown in Fig. \ref{fig:abstractA_Z}, computes 
$f: \mathbb{B}^k \rightarrow \mathbb{B}^k$ and can thus be analyzed as the
function $f: \Fkk \rightarrow \Fkk$. Over $\Fkk$ this function can be 
represented as the polynomial $Z=\Func(A)$. This is trivially generalized
when there are multiple $k$-bit inputs $A_1,A_2,\dots,A_i$, i.e. $Z=\Func(A_1,\dots,A_i)$.
Now assume the word-size of the input differs from the output, that is the circuit
computes $f: \mathbb{B}^m \rightarrow \mathbb{B}^n$ for $m \neq n$. This can 
be represented as a function over Galois fields as $f: \F_{2^m} \rightarrow \F_{2^n}$.
This function can be analyzed over the field $\Fkk$ such that $\Fkk \supset \F_{2^m}$ 
and $\Fkk \supset \F_{2^m}$, where $k=LCM(m,n)$.

\subsection{Dissertation Contributions}
To solve the problem of word-level abstraction, this dissertation proposes 
a full solution consisting of three main contributions.

\begin{enumerate}
\item A theory for finding the word-level abstraction from a bit-level circuit over Galois fields is created.
The given bit-level circuit implementation is modelled as a system of
polynomials over the field.
This theory is derived using techniques from computer-algebra, notably the theory of
\Grobner basis \cite{pruss:iwls13}.
\item Using this theory, new algorithms based on symbolic computation are developed to 
derive the word-level abstraction. The algorithms are designed to be applicable to 
industry-size arithmetic circuits over Galois fields \cite{pruss:dac14}\cite{sun:date15}.
A complexity analysis of the algorithmic approach is also presented. Furthermore, the
approach is also generalized to make it applicable to arbitrary combinational circuits.
Finally, we show how the approach can be used to exploit the hierarchical structure of
large Galois field multipliers designed over composite fields.
\item A custom software tool implementation of the algorithmic approach is described, including
an analysis of efficient data structures designed for this purpose \cite{pruss:tcad15}.
\end{enumerate}

Experiments show that the proposed solution can abstract canonical, word-level, 
polynomial representations of Galois field arithmetic circuits up to $1024$-bits in
size, while other contemporary approaches are infeasible beyond a $32$-bit designs.

%an 
%approach based on computer-algebra, algebraic geometry - notably the theory of
%{\it Gr\"obner bases} \cite{gb_book} \cite{buchberger_thesis}.
%and {\it elimination ideals} \cite{ideals:book}, as our approach for 
%abstracting canonical word-level representations from bit-level Galois 
%field arithmetic circuits. 
%Computer-algebra techniques easily integrate with 
%Galois field theory and allow for polynomial abstraction of the circuit over 
%the field itself. Thus, if the circuit computes a function over some field $
%\mathbb{F}_{2^k}$, the resulting abstraction is a word-level polynomial 
%function over the same field $\mathbb{F}_{2^k}$.

%The given bit-level circuit implementation is first modeled as a system of
%polynomials over the field. Using the theory of elimination ideals, we 
%derive an elimination term ordering and prove that by using this ordering we 
%can obtain a canonical word-level representation of a circuit by 
%computing a Gr\"obner basis of the polynomials. However, complexity 
%of the computation of a Gr\"obner basis proves to be prohibitive for circuits
%of practical size. Therefore, we employ a select sub-set of computations 
%from the Gr\"obner basis theory to overcome this complexity and obtain the
%abstraction. We prove that this simpler computation can be 
%performed via a polynomial reduction process, and engineer a custom 
%verification tool for this purpose. Using this approach, we are able to 
%successfully 
%abstract word-level representations of Galois field arithmetic circuits up 
%to $571$-bits, which is the {\it largest recommended NIST standard} for ECC.

\section{Dissertation Organization}
The rest of this dissertation is organized as follows. Chapter
\ref{ch:prev} reviews previous applicable work and highlights their
drawbacks with respect to the canonical, word-level abstraction problem. 
Chapter \ref{ch:prelim} describes the properties of Galois fields, 
$\mathbb{F}_{2^k}$, and explains the process of constructing them.
It also describes how to design arithmetic circuits over such fields, their 
complexities, and the role of these circuits in Elliptic Curve Cryptography.
Chapter \ref{ch:ideals} provides a theoretical background of 
computer-algebra and Gr\"obner bases and explains their application
to Galois fields. 
Chapter \ref{ch:abstract} describes an approach to abstract 
word-level polynomial representations of combinational circuits using a 
Gr\"obner basis computation.
Chapter \ref{ch:improv} improves on this word-level abstraction approach to 
make it applicable to much larger circuits.
Chapter \ref{ch:generalize} generalizes the abstraction approach to make
it applicable to circuits with varying operand word-lengths. It also describes how the 
approach can take advantage of the hierarchy of arithmetic circuits designed over
composite fields.
Chapter \ref{ch:implement} describes the implementation details of a 
custom abstraction tool and gives experimental results of
abstracting large Galois field multiplier circuits.
Chapter \ref{ch:concl} concludes the dissertation and outlines potential future 
research for continuation of this work.

>>>>

\chapter{Introduction}
\label{ch:intro}
With the increasing size of integrated
circuits, sequential circuit designers face complicated problems of
design errors in specification models and implementations. These
errors are usually modeled as ``bad" states, and the
circuits/functional components are modeled as finite state machines
(FSMs). Once state reachability is analyzed, the existence of errors
can be identified by checking whether ``bad" states are {\it
  reachable} from certain initial states. Temporal logic {\it model
  checking} (MC) formulations and solvers are often used for this
purpose. Once the designs and specification models are validated using
model checking, optimized implementations of sequential circuits are
synthesized. A subsequent problem then needs to be solved to prove
that the sequential circuit implementations are equivalent to the
original specification models, i.e. {\it Sequential Equivalence
  Checking   (SEC)}.

Reachability analysis forms the backbone of most sequential
verification techniques. As the state-space of FSMs increases,
reachability analysis forms a fundamental bottleneck in sequential
verification. Contemporary approaches employ various techniques to
overcome this state-explosion problem: i) Bounded model checking
\cite{bitlevel1} traverses the FSMs for a fixed number of steps $k$
($k$-BMC) to check whether a property violation can occur in $k$ or
fewer steps. ii) Analyze over-approximations (or {\it abstractions})
of the state-space. Abstraction proves properties on the system by
first simplifying it, and when the abstraction does not satisfy the
same properties as the original one, a process of refinement is
needed. {\it Craig interpolants} \cite{craig-interpolate}
\cite{pudlak:ci} are used as a means of abstraction and reachability
analysis for linear temporal logic (LTL) model checking
\cite{mcmillan:interpolation}. Similarly, counterexample guided
abstraction refinement (CEGAR) \cite{cegar-journal} uses proofs of
unsatisfiability (UNSAT cores) to refine the abstractions. iii) the
recent breakthrough method of \cite{bradley:pdr} where the set of
over-approximations to forward reachable states are refined with
inductive constraints -- property directed reachability (PDR).

While the above techniques have made significant strides in sequential
verification, numerous practical instances remain unsolved. One issue
with all of the above techniques is that they mostly use bit-level
constraints to model the transition relations and sets of 
states. We have access to instances from our industrial
collaborators (Calypto Design Systems, now Mentor Graphics) where
the designs are expressed at the level of bit-vector words
(e.g. Matlab code, Verilog), and these word-level abstractions are
rarely exploited in verification. The problem is further exacerbated
when there are arithmetic operators on word-level operands embedded in
the control logic. While attempts have been made towards word-level
predicate abstraction \cite{Jain05wordlevel} \cite{mcmillan:cav06}
\cite{mcmillan:impact}, {\it using a purely word-level representation
  of the state-space, the properties and their abstractions have not
  been fully explored as another dimension in improving sequential
  verification.}

There are several advantages in exploiting word-level information for
verification. A number of designs have  their
datapaths and/or system-level models described as word-level RTL
models.  Exploiting word-level instead of bit-level information is one
way of abstraction -- a key technique to reduce the state space of a
FSM. It has the effect of combining sets of states with similar 
properties. During reachability analysis, if we use bit-level
variables to  represent the states, the representations may become too
large to handle. However, when a ``bundle" of bit-level variables are
represented as {\it only one word-level variable}, the set of
reachable states can be represented by a  word-level constraint
expression; which may lower verification complexity.

{\bf Overall Research Objective:} In this proposal we investigate a
set of new, promising approaches for {\it word-level representation,
reachability analysis and abstraction} with applications to SEC,
$k$-BMC and abstraction-based unbounded LTL model checking
(A-UMC). Our techniques operate at the word-level and they are based
largely on the concepts from {\it algebraic geometry}.

For word-level SEC, we are given two designs, or their corresponding
Mealy/Moore FSMs ${\mathcal{M}}_1,{\mathcal{M}}_2$, along with their
initial starting states $S_0^1,S_0^2$. We wish to prove the absence of
a sequence of inputs (string) that distinguishes the initial
states \cite{coudert:iccad90}
\cite{coudert1990verification}. Fundamentally, this requires 
the construction of a product machine; and the main research problem
relates to that of performing {\it FSM traversal \cite{touati:iccad90}
but  at the word-level}. Analogously, in the case of MC, the problem
is setup w.r.t. a FSM $\mathcal{M}$, a set of initial states $S_0$ and
a set of property states $p$. Techniques are to be researched that
verify that there exist no sequence of transitions from an initial
state to a non-property state (``bad'' state). These problems have to
be solved in the context of word-level verification -- i.e. data
representation, abstraction (Craig interpolation, UNSAT cores) and
algorithm execution has to be carried out at word-level.

{\bf The Algebraic Geometry Model and Rationale:} We will represent
the FSMs -- the transition relations -- by means of a set of
multi-variate polynomials with coefficients from the finite (Galois)
field $\Fkk$ of $2^k$ elements, i.e. polynomials in the ring
$\Fkk[x_1,\dots,x_d]$. Each state of a FSM is identified with a
Boolean assignment to a set of $k$-bit state register variables
$S=\{s_0,\dots,s_{k-1}\}$. Therefore, we can consider each ($k$-bit)
state as a word-level element $S$ of the finite field
$\Fkk$. Algorithms can directly operate on polynomials in word-level
variable $S$.

Boolean functions with $k$-bit inputs and $k$-bit outputs 
$f: \B^k \rightarrow \B^k,\B = \{0, 1\}$ can be construed as functions
$f: \Fkk \rightarrow \Fkk$. It is well-known that over the finite
field ($\Fq$) of  $q$ elements, {\it every function} $f: \Fq
\rightarrow \Fq$ is a polynomial function \cite{ff:1997}. Moreover,
there exists a unique canonical polynomial $\FF$ that describes $f$.
This implies that one can derive a {\it  canonical, polynomial
  abstraction} of the function as $Z = \FF(A)$ where $Z, A$ are
word-level symbols representing $k$-bit operands. The concept also
generalizes to functions with different input/output bit-vector sizes,
i.e. functions of the type $f: \B^n \rightarrow \B^m$, modeled as
polynomials over $f:{\mathbb{F}}_{2^k} \rightarrow
{\mathbb{F}}_{2^k}$, where $k=LCM(n,m)$ \cite{ff:1997}.  
This implies that the FSM's transition relations can be
represented as polynomial functions (ideals) in $\Fkk$, and values of
state variables can be represented as solutions to these polynomials
(variety of the ideal). Subsequently, the ideal-variety correspondences
in algebraic geometry can be applied to implement symbolic reasoning
about state reachability. Moreover, as ${\mathbb{F}}_2 \subset \Fkk$,
our model provides a single, unified and bit-precise representation
for both bit-level (${\mathbb{F}}_2$) and word-level ($\Fkk$)
constraints.

The decision and abstraction procedures in our setting will rely on
the theory and technology of {\it \Grobner bases (GB)}. GB-based
algebraic reasoning is very powerful; in fact it is known to be
strictly stronger than resolution \cite{CEI:stoc-96}. Therefore, in
light of the above discussion, using concepts from algebraic geometry
and \Grobner bases over $\Fkk$, we can introduce another dimension of
word-level abstraction to the techniques in sequential verification.

{\bf The core focus of this research:} The use of algebraic geometry
for sequential circuit verification and symbolic model checking has
been presented before. Avrunin presented the
concept of symbolic MC using algebraic geometry in
\cite{Avrunin:CAV}. Later, in \cite{vardi-iasted07}, Vardi presented
GB-algorithms for CTL, LTL and 
bounded MC over Boolean rings. However, these approaches are a
straight-forward transformation of the problem to {\it bit-level}
Boolean GB engines which are used in lieu of BDDs or SAT solvers. All
the concepts of word-level reachability, abstraction-refinement using
interpolation or UNSAT cores, etc., that we desire was not the focus of
\cite{Avrunin:CAV} \cite{vardi-iasted07}.

In this research, we will focus on three specific research problems: 
\begin{enumerate}
\item How to perform reachability analysis at the word-level? This
  of course forms the foundation for word-level SEC and MC.
\item What is the analog of Craig interpolation in an algebraic
  geometry setting (theoretical challenge), and how can we compute
  various interpolants using GB-engines efficiently (practical
  implementation)? This will enable word-level abstraction of the
  state-space. 
\item Explore the concept, and the computation, of unsatisfiable (UNSAT)
  cores of  a set of polynomials. Many abstraction-refinement
  techniques make use of UNSAT cores of CNF-formulas in MC. Therefore,
  our investigation of UNSAT cores of polynomial ideals will enable
  similar applications at word-level. 
\end{enumerate}

\section{Design, Test and Verification of Modern Sequential Circuit Components}\section{Hardware Design and Verification Overview}
The typical design flow of a hardware system, 
as shown in Fig. \ref{fig:cadflow}, starts with a hardware system 
specification, which describes the necessary functions and parameters that 
the system must perform and adhere to. The specification is typically
modelled using a transaction-level model (\emph{TLM}), which describes 
communication details between large circuit modules. 
The TLM is then translated 
into a register-transfer-level (\emph{RTL}) description, which is composed of abstracted, 
interconnected circuit blocks that compose the entire system. RTL is 
typically implemented 
in hardware description languages (\emph{HDL}) such as Verilog and VHDL, 
which are the most popular choices in the industry.
Next, the RTL is optimized and converted into a \emph{netlist}, i.e. a large 
collection of small physical blocks (MOSFET, Boolean logic gates, etc.) and 
the inter-connections (wires) between them. 
Lastly, the netlist is further optimized and then mapped onto a physical space 
on a chip, which is then sent off for fabrication. This entire design flow
is automated by Computer-Aided Design (\emph{CAD}) tools.

@@

\section{Objective, Motivation and Contribution of this Dissertation}
\subsection{Word-level Reachability and FSM Traversal}
\subsection{Application to Galois Field Multipliers}
\subsection{Abstractions for Sequential Circuits}
\subsection{Other Applications}

\section{Dissertation Organization}
{\bf Proposal Organization:} The rest of the proposal is organized as
follows: The following Subsection describes the intellectual merit of
the proposed work.  Section \ref{sec:prelim} covers preliminary
concepts and notation. Section \ref{sec:reach} describes the basic
concept of word-level FSM traversal and the research problems to
investigate. Section \ref{sec:craig} explores the algebraic geometry
analog of Craig Interpolants. Section \ref{sec:core} describes
algorithmic techniques to derive UNSAT cores of polynomial
ideals. Section \ref{sec:plan} puts the whole research in perpective,
by demonstrating  with the help of an example how abstraction via
unsat cores in algebraic geometry can simplify BMC. 
Section \ref{sec:impact} addresses the broader impact of the
proposed activity and concludes the proposal.
